### 2025-02-18 paper.tex
Add citations

\documentclass[runningheads]{llncs}

% TODO: standardise indices vs indexes
% TODO: standardise dots after equations
% TODO: rename variable maps to variable indices everywhere
% TODO: Title Case for paragraphs

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[british]{babel}
\usepackage{newunicodechar}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsmath,amssymb}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{hyperref}
\usepackage{csquotes}
\usepackage[noend]{algorithm2e}
\usepackage{siunitx}
\usepackage{caption}
\usepackage{subcaption}

\bibliographystyle{splncs04}

\renewcommand\UrlFont{\color{blue}\rmfamily}
\urlstyle{rm}
\def\orcid#1{{\href{http://orcid.org/#1}{\protect\raisebox{-1.25pt}{\protect\includegraphics{orcid.pdf}}}}}
\def\showpgfcircle{\tikz[baseline=-0.7ex]\node[mark size=0.7ex]
{\pgfuseplotmark{o}};}
\def\showpgfsquare{\tikz[baseline=-0.7ex]\node[mark size=0.7ex]
{\pgfuseplotmark{square}};}
\SetKw{KwIn}{in}
\SetKw{KwLet}{let}
\SetKw{KwYield}{yield}
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output}
\SetKwProg{Procedure}{Procedure}{}{}

\newunicodechar{₀}{\ensuremath{_{0}}}
\newunicodechar{₁}{\ensuremath{_{1}}}
\newunicodechar{₂}{\ensuremath{_{2}}}
\newunicodechar{₃}{\ensuremath{_{3}}}
\newunicodechar{₄}{\ensuremath{_{4}}}
\newunicodechar{₅}{\ensuremath{_{5}}}
\newunicodechar{₆}{\ensuremath{_{6}}}
\newunicodechar{₇}{\ensuremath{_{7}}}
\newunicodechar{₈}{\ensuremath{_{8}}}
\newunicodechar{₉}{\ensuremath{_{9}}}
\newunicodechar{ₙ}{\ensuremath{_{n}}}
\newunicodechar{ₘ}{\ensuremath{_{m}}}
\newunicodechar{ᵢ}{\ensuremath{_{i}}}
\newunicodechar{ⱼ}{\ensuremath{_{j}}}
\newunicodechar{ₗ}{\ensuremath{_{l}}}
\newunicodechar{ₖ}{\ensuremath{_{k}}}
\newunicodechar{α}{\ensuremath{\alpha}}
\newunicodechar{σ}{\ensuremath{\sigma}}
\newunicodechar{τ}{\ensuremath{\tau}}
\newunicodechar{λ}{\ensuremath{\lambda}}
\newunicodechar{μ}{\ensuremath{\mu}}
\newunicodechar{η}{\ensuremath{\eta}}
\newunicodechar{Γ}{\ensuremath{\Gamma}}
\newunicodechar{Δ}{\ensuremath{\Delta}}
\newunicodechar{Φ}{\ensuremath{\Phi}}
\newunicodechar{Ψ}{\ensuremath{\Psi}}
\newunicodechar{Ω}{\ensuremath{\Omega}}
\newunicodechar{∈}{\ensuremath{\in}}
\newunicodechar{∉}{\ensuremath{\notin}}
\newunicodechar{≡}{\ensuremath{\equiv}}
\newunicodechar{≢}{\ensuremath{\nequiv}}
\newunicodechar{≠}{\ensuremath{\neq}}
\newunicodechar{≤}{\ensuremath{\leq}}
\newunicodechar{≥}{\ensuremath{\geq}}
\newunicodechar{≺}{\ensuremath{\prec}}
\newunicodechar{≔}{\ensuremath{\coloneq}}
\newunicodechar{∧}{\ensuremath{\land}}
\newunicodechar{∨}{\ensuremath{\lor}}
\newunicodechar{∪}{\ensuremath{\cup}}
\newunicodechar{∩}{\ensuremath{\cap}}
\newunicodechar{⊆}{\ensuremath{\subseteq}}
\newunicodechar{⋃}{\ensuremath{\bigcup}}
\newunicodechar{→}{\ensuremath{\rightarrow}}
\newunicodechar{↔}{\ensuremath{\leftrightarrow}}
\newunicodechar{↦}{\ensuremath{\mapsto}}
\newunicodechar{⊢}{\ensuremath{\vdash}}
\newunicodechar{∅}{\ensuremath{\emptyset}}
\newunicodechar{∷}{\ensuremath{\mathbin{::}}}
\newunicodechar{⇒}{\ensuremath{\Rightarrow}}
\newunicodechar{ℕ}{\ensuremath{\mathbb{N}}}
\newunicodechar{⟨}{\ensuremath{\langle}}
\newunicodechar{⟩}{\ensuremath{\rangle}}
\newunicodechar{⊤}{\ensuremath{\top}}
\newunicodechar{⊥}{\ensuremath{\bot}}
\newunicodechar{×}{\ensuremath{\times}}

\newtheorem{lem}{Lemma}

\setminted[lean4]{extrakeywords={aesop cases add aesop? intro simp simp_all only split apply on_goal next rename_i safe unsafe norm constructors forward destruct norm_num done add_aesop_rules rfl subst ext}}
\newmintinline[lean]{lean4}{bgcolor={},ignorelexererrors=true}
\newminted[leancode]{lean4}{bgcolor={},ignorelexererrors=true,fontsize=\footnotesize,autogobble}
\BeforeBeginEnvironment{leancode}{\begin{tcolorbox}}
\AfterEndEnvironment{leancode}{\end{tcolorbox}}
\usemintedstyle{xcode}

\newcommand{\xcom}[1]{{\textcolor{cyan}{Xavier: #1}} }
\newcommand{\jcom}[1]{{\textcolor{orange}{Jannis: #1}} }

\newcommand{\Lam}[2]{\ensuremath{\lambda\, #1,\; #2}}
\newcommand{\All}[2]{\ensuremath{\forall\, #1,\; #2}}
\newcommand{\mvar}[1]{\ensuremath{?#1}}
\newcommand{\Prop}{\ensuremath{\mathrm{Prop}}}
\newcommand{\vars}{\ensuremath{\mathrm{vars}}}
\newcommand{\dom}{\ensuremath{\mathrm{dom}}}
\newcommand{\sub}{\ensuremath{\mathrm{sub}}}
\newcommand{\lvl}{\ensuremath{\mathrm{lvl}}}
\newcommand{\Rules}{\ensuremath{\mathcal{R}}}
\newcommand{\Hyps}{\ensuremath{\mathcal{H}}}
\newcommand{\States}{\ensuremath{\mathcal{S}}}
\newcommand{\CMatches}{\ensuremath{\mathcal{C}}}
\newcommand{\addHyp}{\ensuremath{\mathrm{addHyp}}}
\newcommand{\delHyp}{\ensuremath{\mathrm{delHyp}}}
\newcommand{\addMatch}{\ensuremath{\mathrm{addMatch}}}
\newcommand{\powerset}{\ensuremath{\mathcal{P}}}

\begin{document}

\title{Incremental Forward Reasoning for White-Box~Proof~Search}

%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here

\author{Jannis Limperg\orcid{0000-0002-8861-5231} \and
  Xavier Généreux\orcid{0000-0003-4952-9557}}

\authorrunning{J.~Limperg and X.~Généreux}
\institute{Ludwig-Maximilians-Universität München, Munich, Germany\\
\email{\{jannis.limperg,xavier.genereux\}@lmu.de}}

\maketitle

\begin{abstract}
  Several proof assistants provide automation tactics based on tableau-style tree search, such as Isabelle's and Rocq's auto and Lean's Aesop.
  In this setting we consider \emph{forward rules}, which apply a given theorem, say, $A → B → C$, to any goal containing hypotheses $A$ and $B$, adding $C$ as a new hypothesis.
  When treated naively, such rules are tried on every goal encountered during the search, leading to repeated unifications of premises $A$ and $B$ with the hypotheses of each goal.
  We present an approach to forward rules that avoids some of this repeated work by taking advantage of similarities between successive goals.
  For each goal, we cache partial applications of forward rules in a custom data structure, the \emph{forward state}, which enables efficient updates.
  Our technique is compatible with any search strategy and most logics.
  It has been implemented in Aesop.

  \keywords{Forward Reasoning \and Forward Chaining \and Tactics \and Interactive Theorem Proving.}
\end{abstract}

\section{Introduction}

In many proof assistants, simple \enquote*{white-box} automation tactics based on tableau-style tree search play a substantial role in practice: Isabelle's~\cite{Isabelle} and Rocq's~\cite{Coq} auto~\cite{IsabelleAuto}, ACL2's \enquote{waterfall}~\cite{ACL2}, various PVS~\cite{PVS,PVS-tutorial} tactics, Lean's~\cite{Lean4} Aesop~\cite{Aesop}, etc.
These tactics operate on \emph{goals} $Γ ⊢ A$, where the \emph{context} $Γ$ is a list of assumptions and $A$ is the proposition to be proved.
Faced with an initial goal, the tactics try various user-specified \emph{rules}, which correspond to admissible inferences in the prover's logic.
A successful rule reduces the current goal to zero or more subgoals, which are solved recursively until either a full proof is found or no rule applies to any open goal or the search times out.

Typical rules would be the $∧$-introduction lemma $A → B → A ∧ B$, which, when applied backwards, reduces a goal $Γ ⊢ A ∧ B$ to subgoals $Γ ⊢ A$ and $Γ ⊢ B$; or left $∨$-introduction, which reduces $Γ ⊢ A ∨ B$ to $Γ ⊢ A$.
Such \emph{backward rules} are the most common rule type, but most search tactics allow users to use any tactic as a rule.
These rule tactics can output arbitrary subgoals, subject only to the condition that proofs of the returned subgoals can be assembled into a proof of the goal on which the rule was run.
\jcom{The framing here is wrong: we want to talk about non-forward rules that mess with the context, e.g. case splitting.}

ACL2, PVS and Aesop provide special support for \emph{forward rules}.
Given, for example, the transitivity lemma $\All{x~y~z}{x < y → y < z → x < z}$, a forward rule run on the goal $h₁ : a < b,\, h₂ : b < c ⊢ A$ produces the single subgoal $h₁ : a < b,\, h₂ : b < c,\, h₃ : a < c ⊢ A$.
Hence forward rules can be used to establish additional facts that may be relevant for other rules.
For example, the Aeneas project~\cite{Aeneas} uses Aesop's forward rules \jcom{ensure that this is accurate} in its proof automation to establish facts that are later used by an arithmetic decision procedure.

However, Aesop's support for forward rules was, until recently, quite basic.
In particular, each forward rule was tried on each goal separately (subject to pre-filtering by a discrimination tree~\cite{DiscriminationTrees} index).
This is wasteful since in practice, many rules produce subgoals with a similar---or even the same---context as the initial goal.
For example, the ∧-introduction rule affects only the goal's target, not its context.
Still, all forward rules applied to the initial $Γ ⊢ A ∧ B$ would be tried again on the subgoals $Γ ⊢ A$ and $Γ ⊢ B$, with predictably little effect.

In this paper, we present an alternative way to apply forward rules that addresses this issue.
The central idea is to remember, for each goal $Γ ⊢ A$, which forward rules can be partially applied to which sets of hypotheses.
When a non-forward rule then generates subgoals $Δ₁ ⊢ A₁$, \dots, $Δₙ ⊢ Aₙ$, the new contexts $Δᵢ$ are likely similar to $Γ$.
Hence it is generally cheaper (and often much cheaper) to start with the partial matches for $Γ$ and update them for each of the $Δᵢ$.
The update first deletes partial matches that include hypotheses no longer present in $Δᵢ$.
Afterwards it adds new partial matches for the hypotheses that are new in $Δᵢ$, extending the existing partial matches for a rule if the new hypothesis is compatible with them.
Whenever this process produces a full match---i.e., all premises of a forward rule are satisfied by a set of hypothesis in $Δᵢ$---the rule is applied and adds a corresponding new hypothesis to $Δᵢ$.
Our main contribution is a data structure, the \emph{forward state}, that organises the partial matches in such a way that the update can be done efficiently.

In addition to this \emph{incrementality}, the forward state satisfies two other desiderata:
\begin{itemize}
  \item It is \emph{persistent}, i.e.\ forward states for parent and child goals share much of their structure.
        This is desirable since Aesop supports search strategies other than depth-first search and must therefore keep all goals encountered during the search, along with their forward states, in memory.
  \item It reduces the number of \emph{unifications} performed when matching hypotheses against rule premises.
        In dependent type theory, unification happens up to definitional equality, which makes it quite expensive.
        We therefore want to avoid redundant unifications.
\end{itemize}

Our evaluation (Sec.~\ref{sec:evaluation}) shows that this new implementation of forward reasoning in Aesop substantially outperforms the previous, naive implementation on some (though not all) synthetic benchmarks.
The implementation, benchmarks and evaluation data are available in a supplement to this paper~\cite{supplement}.
Mainline Aesop already uses the new implementation by default (TODO true?).

\section{Preliminaries}

We first introduce some terminology for matches and for forward rules (which we identify with their underlying lemmas).
Our foundation in this paper is dependent type theory, but our technique can be adapted to other logics.

Given a lemma $\All{(x₁ : A₁) \dots (xₙ : Aₙ)}{B}$, the $xᵢ$ are \emph{premises} and $B$ is the lemma's \emph{conclusion}.
The type $Aᵢ$ of the $i$-th premise may depend on (i.e., refer to) any premise $xⱼ$ with $j < i$.
Implication $A → B$ is an abbreviation for $\All{x : A}{B}$ where $B$ does not depend on $x$.
Propositions are (certain) types and we do not distinguish between the two concepts.

Premises on which a later premise depends are \emph{variables}.
All other premises are \emph{input premises}.
For example, the lemma $\All{x~y~z}{x < y → y < z → x < z}$ has variables $x, y, z$ and (unnamed) input premises of types $x < y$ and $y < z$.
If a variable occurs in at least two premises, like $y$ above, it is \emph{shared}.

For a lemma with $n$ input premises, we assign to each input premise a unique number between $1$ and $n$.
This is the premise's \emph{slot}, which we often identify with the premise itself.
For now, we assume that the input premises are numbered consecutively, so in the above example slot~1 has type $x < y$ and slot~2 has type $y < z$.
The \emph{shared variables of slot $i$}, $\vars(i)$ are those variables that appear in the type of $i$ and also in at least one type of a slot $j < i$.

If $h : A$ is a hypothesis in the context of a goal and $A$ unifies with the type $B$ of slot $i$, we say that $h$ \emph{matches} $i$.
Unification induces a substitution $σ$ such that $A[σ]$ is definitionally equal to $B[σ]$; we call this substitution $\sub_{i}(h)$.
For now we do not specify exactly what \enquote{unifies} means; see Sec.~\ref{sec:implementation}.

A \textit{match} $m$ for a give rule $r$ with $n$ slots is a list containing at least one and at most $n$ hypotheses.
We interpret $m$ as a partial map from slots of $r$ to hypotheses, so $\dom(m) = \{i \mid 1 ≤ i ≤ |m| \}$ and for each slot $i ∈ \dom(m)$, $m(i)$ is the hypothesis assigned to $i$.
Each match $m$ must satisfy the following requirements:
\begin{enumerate}
  \item\label{cond:match:matches} For each slot $i ∈ \dom(m)$, the hypothesis $m(i)$ matches $i$.
  \item\label{cond:match:compat} For all slots $i,j ∈ \dom(m)$, the substitutions $\sub_{i}(m(i))$ and $\sub_{j}(m(j))$ are \emph{compatible}.
        This means that the substitutions agree on all common variables, i.e.\ for each variable $x ∈ \dom(\sub_{i}(m(i))) ∩ \dom(\sub_{j}(m(j)))$ we have $\sub_{i}(m(i))(x) = \sub_{j}(m(j))(x)$.
\end{enumerate}
The \emph{substitution of a match $m$}, $\sub(m)$, is the union of the substitutions induced by the hypotheses in $m$, i.e. $\sub(m) ≔ ⋃_{i ∈ \dom(m)}\sub_{i}(m(i))$.
This is well-defined by requirement~\ref{cond:match:compat}.
The \textit{level of a match $m$}, $\lvl(m)$, is the length of $m$.
A match for rule $r$ is \textit{complete} if $i ∈ \dom(M)$ for every slot $i$ of $r$, and \emph{partial} otherwise.

A partial match $m$ for rule $r$ corresponds to a partial application of the rule to the terms given in $\sub(M)$.
The application is type-correct by requirements~\ref{cond:match:matches} and~\ref{cond:match:compat}.
If $m$ is complete, then $\sub(m)$ contains a term for every slot and every variable---hence for every premise---of $r$ and we have a full application.

Note that matches represent only those partial applications where arguments are added without gaps.
Hence for a rule $r : A → B → C → D$, the partial applications $r~a$ and $r~a~b$ for some $a : A$ and $b : B$ correspond to matches, but not $r~\text{--}~b$.
This will prevent us from constructing $r~a~b$ in two different but ultimately redundant ways.
The term orders used in various saturation calculi serve a similar function.

\section{Naive Forward Reasoning}
\label{sec:naive}

\jcom{Formulate this in terms of matches.}
Prior to this work, Aesop used what we call naive forward reasoning~\cite{Aesop}.
This approach treats forward rules essentially like any other Aesop rule.
When Aesop selects a forward rule $r : \All{(x₁ : A₁) \dots (xₘ : Aₘ)}{B}$ as the highest-priority rule currently applicable to a goal $Γ ⊢ T$, we first unify the last premise $Aₘ$ with each hypothesis $h : C$ such that $C$ and $Aₘ$ likely unify (as determined by a discrimination tree index).
If successful, this results in a substitution $σ₁$ containing $xₘ$ and any variables occurring in $Aₘ$.
We then unify $A_{m-1}[σ₁]$ with the type of each hypothesis in the context, yielding a substitution $σ₂$, and proceed with $A_{m-2}[σ₂]$, etc., until we obtain a complete match for $r$ (skipping premises $A_{m-i}$ if they are already determined by $σᵢ$).
If a premise $A_{m-i}[σᵢ]$ does not unify with any hypothesis type, we backtrack and consider other potential matching hypotheses for $A_{i+1}$.
These may lead to a different substitution $τ$ and therefore to matching hypotheses for $A_{i}[τ]$.
If this process yields at least one full application $r~a₁ \dots aₘ$, we can add a new hypothesis of type $T = B[x₁ ↦ a₁, \dots, xₘ ↦ aₘ]$, meaning we generate the subgoal $Γ, h : T$ (unless $Γ$ already contains a hypothesis of type $T$).

This approach is efficient when the last premise of a rule is not present in the context since the rule is then excluded by indexing and never run.
If the last premise is present, however, we perform a large number of unifications between premises and hypotheses (though some of these could perhaps be avoided heuristically~\cite{TammetSubsumption}).
Furthermore, the whole process is repeated when we proceed to another goal with a similar context.
In particular, the contexts of subgoals produced by whatever rule is eventually applied to the current goal are likely to be substantially similar, if not identical, to the current context, but these similarities are not exploited by the naive algorithm.

\section{Incremental Forward Reasoning}%
\label{sec:incremental}

With the weaknesses of the naive approach in mind, we introduce an incremental, stateful solution to the problem of finding complete matches.
The main idea is to store, for each goal $Γ ⊢ A$, the partial matches implied by the hypotheses in $Γ$, and to reuse these partial matches for subgoals $Δ ⊢ B$.
This rests on the assumption that $Δ$ is likely similar to $Γ$ in practice, and that it is therefore cheaper to update the partial matches for $Γ$ than to compute those of $Δ$ from scratch.
Our main contribution is a data structure, the \emph{forward state}, that allows us to perform these updates efficiently.

\paragraph{Forward State.}
We associate to each goal in Aesop's search tree a forward state that reflects the goal's context.
Let $\States$ be the set of forward states, $\Hyps$ the set of hypotheses, $\Rules$ the set of rules, $\CMatches$ the set of complete matches and $\powerset(X)$ the powerset of $X$.
The forward state's interface then consists of two functions
\begin{equation*}
  \addHyp : \States → \Hyps → \States × \powerset(\Rules × \CMatches) \qquad \delHyp : \States → \Hyps → \States
\end{equation*}
The $\addHyp$ function adds a hypothesis to the forward state.
Hence, if $s$ is the forward state corresponding to context $Γ$ and $\addHyp(s, h : T) = (s', C)$, then $s'$ is the forward state corresponding to $Γ,\, h : T$.
Similarly, if $s$ is the forward state corresponding to $Γ, h : T$, then $\delHyp(s, h : T)$ is the forward state corresponding to $Γ$.
The second output of $\addHyp$, $C$, is a set of pairs $(r, m)$ where $r$ is a rule and $m$ is a complete match for $r$.
These are the matches that were completed by adding $h : T$.

When Aesop is run on an initial goal with context $h₁ : T₁, \dots, hₙ : Tₙ$, the forward state for it is built by starting with an empty state and applying $\addHyp$ to each hypothesis $hᵢ$.
Afterwards, whenever an Aesop rule is run on a goal with context $Γ$, producing subgoals with contexts $Δ₁, \dots, Δₙ$, we derive the forward state for each subgoal context $Δⱼ$ from the forward state for $Γ$, in two steps:
\begin{enumerate}
  \item Apply $\delHyp$ once for each hypothesis that was deleted, i.e.\ that appears in $Γ$ but not in $Δⱼ$.
  \item Apply $\addHyp$ once for each hypothesis that was added, i.e.\ that appears in $Δⱼ$ but not in $Γ$.
\end{enumerate}
Changes to a hypothesis (e.g.\ a change of its type) are treated as a deletion followed by an addition.

\paragraph{Rule State.}
Partial matches are specific to each rule, so the forward state naturally decomposes into a set of \emph{rule states}, one per rule.
Given a goal $Γ ⊢ A$ and a rule $r$, the rule state for $r$ contains exactly the partial matches of $r$ in context $Γ$.
A rule state supports $\addHyp$ and $\delHyp$ methods analogous to those on forward states.
Adding a hypothesis to a forward state then amounts to adding it to each rule state (which may yield new complete matches for the rule); deleting a hypothesis amounts to deleting it from each rule state.
We initially focus on how to add hypotheses, since this question motivates much of the design of the rule state data structure.

When a hypothesis is added to a rule state, two things can happen.
First, the hypothesis may allow us to extend partial matches already present in the state.
For instance, consider the rule $r : \All{x}{P~x → Q~x → R~x → S~x}$ and a rule state containing the partial match $[h₁]$, which assigns some hypothesis $h₁ : P~a$ to the first slot of type $P~x$.
Adding a new hypothesis $h₂ : Q~a$ to this rule state means that we should add to the state the extended match $[h₁, h₂]$.

Second, adding a match may trigger the insertion of further matches.
Suppose that before adding $h₁$ and $h₂$, we had already added a hypothesis $h₀ : R~a$ to the rule state.
Once the partial match $[h₁, h₂]$ arrives, we should recognise that it can be extended with $h₀$ to obtain the (complete) match $[h₁, h₂, h₀]$.

\paragraph{Variable Indices.}
Extending the matches above was possible only because all hypotheses are compatible on the shared variable $x$, i.e.\ they all instantiate $x$ with the same term $a$.
In general, to extend a match $m$ with a hypothesis $h : T$ at slot $i$, we need to ensure that the substitution of $m$, $\sub(m)$, is compatible with the substitution of $h$, $\sub_{i}(h)$.

The rule state therefore uses two \emph{variable indices} $H$ and $M$ to index hypotheses and matches by their instantiations for shared variables, allowing us to quickly determine compatible pairs of matches and hypotheses.
We could use a substitution tree~\cite{SubstitutionTrees}, but our custom data structure is simpler and perhaps more efficient.
The \emph{hypothesis index} $H$ maps each triple $(x, i, t)$ (where $i > 1$ is a slot, $x ∈ \vars(i)$ a variable and $t$ a term) to the set of hypotheses $h$ added to the rule state such that $h$ matches $i$ and $\sub_{i}(h)(x) = t$.
In other words, $H$ indexes the known hypotheses by the slots they match and the instantiations of shared variables.
Note that hypotheses matching slot $1$ are not stored since they can never extend a match.

Similarly, the \emph{match index} $M$ maps each triple $(x, i, t)$ to the set of partial matches $m$ (that can be derived from the hypotheses added to the rule state) such that $i = \lvl(m) $, $x ∈ \vars(i + 1)$ and $\sub(m)(x) = t$.
Matches for the maximal slot are not stored because, being complete, they never need to be extended.
We write $M_{x}(i, t)$ instead of $M(x, i, t)$ and $H_{x}(i, t)$ instead of $H(x, i, t)$.

Given a hypothesis $h : T$ that matches slot $i > 1$ with substitution $σ = \sub_{i}(h)$, the partial matches that can be extended by $h$ are those in slot $i - 1$ whose substitutions are compatible with $σ$ on the variables $\vars(i)$ (i.e., the variables shared between $i$ and any earlier slot).
Assuming for the moment that $\vars(i)$ is nonempty for all slots $i > 1$, these matches are
\[
  \bigcap_{x \in \vars(i)} M_{x}(i-1, σ(x))
\]
Similarly, given a partial match $m$ at level $i$ with substitution $σ = \sub(m)$, the hypotheses that could extend $m$ are
\[
  \bigcap_{x \in \vars(i+1)} H_{x}(i+1, σ(x))
\]
The number of shared variables tends to be small in practice, so these intersections can be computed efficiently.

From these observations we can directly derive two procedures for adding hypotheses and matches to the rule state.
The procedures maintain the invariants of the variable indices while reporting any complete matches discovered during this process.

\begin{algorithm}
  \SetKwInOut{Precondition}{Precondition}
  \SetKwInOut{Postcondition}{Postcondition}
  \Precondition{$H$ and $M$ contain the hypotheses and matches corresponding to a context $Γ$}
  \Input{A hypothesis $h : T$ matching slot $i$ of $r$ with substitution $σ$}
  \Output{Complete matches of $r$ in context $Γ,\, h : T$ that contain $h$}
  \Postcondition{$H$ and $M$ contain the hypotheses and matches corresponding to $Γ,\, h : T$}
  \BlankLine
  \Procedure{$\mathrm{addHyp}(h : T)$}{
    \eIf{$i = 1$}{
      $\mathrm{addMatch}([h])$
    }{
      \For{$x$ \KwIn $\vars(i)$}{
        Insert $h$ into the set $H_{x}(i, σ(h))$
      }
      \KwLet $\mathit{ms} ≔ \bigcap_{x \in \vars(i)} M_{x}(i-1, σ(x))$ \\
      $\mathit{completeMs} \gets ∅$ \\
      \For{$m$ \KwIn $\mathit{ms}$}{
        $\mathit{completeMs} \gets \mathit{completeMs} ∪ \mathrm{addMatch}(m ++ [h])$
      }
      \Return{$\mathit{completeMs}$}
    }
  }
  \BlankLine % without these, the caption is not offset properly
  \BlankLine
  \caption{Adding a hypothesis to the rule state for a rule $r$}%
  \label{alg:addhyp}
\end{algorithm}

The rule state's $\addHyp$ method (Algorithm~\ref{alg:addhyp}) adds a hypothesis $h : T$ matching slot $i$ to the rule state.
(If a hypothesis matches multiple slots, $\addHyp$ is called multiple times.)
For $i = 1$, we add the singleton match $[h]$ to the rule state.
For $i > 1$, we first insert $h$ into the hypothesis index $H$.
Then we query the match index $M$ for matches that can be extended by $h$ and add the extended matches to the rule state.
While adding matches, we may discover new complete matches, which are collected and returned.

% TODO pre- and postcondition?
\begin{algorithm}
  \Input{A partial match $m$ of $r$ at level $i$ with substitution $σ$}
  \Output{A set of complete matches of $r$}
  \BlankLine
  \Procedure{$\mathrm{addMatch}(h : T)$}{
    \eIf{$i$ is complete}{
      \Return{$\{m\}$}
    }{
      \For{$x$ \KwIn $\vars(i + 1)$}{
        Insert $m$ into the set $M_{x}(i, σ(x))$
      }
      \KwLet $\mathit{hs} ≔ \bigcap_{x \in \vars(i + 1)} H_{x}(i + 1, σ(x))$ \\
      $\mathit{completeMs} \gets ∅$ \\
      \For{$h$ \KwIn $\mathit{hs}$}{
        $\mathit{completeMs} \gets \mathit{completeMs} ∪ \mathrm{addMatch}(m ++ [h])$
      }
      \Return{$\mathit{completeMs}$}
    }
  }
  \BlankLine % without these, the caption is not offset properly
  \BlankLine
  \caption{Adding a match to the rule state for a rule $r$}%
  \label{alg:addmatch}
\end{algorithm}

The rule state's $\addMatch$ method (Algorithm~\ref{alg:addmatch}) adds a match $m$ with level $i = \lvl(m)$ to the rule state.
If $m$ is already complete, it is returned.
Otherwise $m$ is inserted into $M$ and we query $H$ for hypotheses that can extend $m$.
Any extended matches thus produced are added to the rule state recursively.

\paragraph{Variable Clusters.}
We have so far assumed that for each slot $i > 1$, the set of shared variables $\vars(i)$ is nonempty, i.e.\ each slot shares at least one variable with at least one of the previous slots.
This assumption is violated for rules such as $\All{x~y}{P~x → Q~y → R~x~y}$: slot~2 ($Q~y$) shares no variables with slot~1 ($P~x$).
As a result, our above algorithms fail.
Suppose we are adding the hypothesis $h₂ : Q~b$, having already added the hypothesis $h₁ : P~a$ and hence the partial match $[h₁]$.
The set of matches compatible with $h₂$ is supposedly $\bigcap_{x \in \vars(2)} M_{x}(i-1, σ(x)) = ∅$, but the match $[h₁]$ can be extended with $h₂$.
In fact, any match at level~1 can be extended with any hypothesis matching slot~2 since the two slots are independent.

To account for this fact, we partition the set of premises of each rule into maximal subsets of premises connected by shared variables, called \emph{variable clusters}, using the same construction as Aesop's \emph{metavariable clusters}~\cite{Aesop}.
The variable clusters share no variables with each other, so each can be treated as a separate rule with its own rule state.
Finding a complete match for a rule amounts to finding one complete match for each of its clusters.
The slots within each cluster can be ordered to ensure that each slot shares at least one variable with the previous slots, satisfying our assumption.

\paragraph{Deletion.}
When a hypothesis $h$ is removed from a goal's context, we must update the forward state accordingly.
This means deleting $h$ from any rule state to which it was previously added, i.e.\ from any rule state for a rule $r$ with a slot $i$ matching $h$.
To facilitate deletion, we cache this information when $h$ is added to the rule state.
When $h$ is deleted, we can then efficiently determine the relevant rule states and delete $h$ from the hypothesis index $H$ at slot $i$ and from the match index $M$ at all slots $j ≥ i$.

\paragraph{Example 1.}
As a first example, consider a rule state for the transitivity rule $r₁ : \All{x~y~z}{x ≤ y → y ≤ z → x ≤ z}$ to which we add, in order, the hypotheses
$h₁ : b ≤ c$ and $h₂ : a ≤ b$.
The rule has two slots of types $x ≤ y$ and $y ≤ z$ which share the variable $y$.

The hypothesis $h₁ : b ≤ c$ unifies with both slots, so we run Algorithm~\ref{alg:addhyp} twice.
For slot 1, unification yields the substitution $\{x ↦ b, y ↦ c\}$.
Since this is the first slot, Algorithm~\ref{alg:addmatch} inserts the new partial match $[h₁]$ into the set $M_{y}(1, c)$.
This match cannot be extended since the hypothesis index $H$ does not contain a hypothesis for slot 2.

For slot 2, unification yields the substitution $σ ≔ \{y ↦ b, z ↦ c\}$, so we insert $h₁$ into $H_{y}(2, b)$.
No matches at level~1 can be extended by $h₁$ since
\[
  \bigcap_{x \in \vars(2)} M_{x}(1, σ(x)) = M_{y}(1, b) = \emptyset
\]
At this point, $H$ and $M$ are in the state shown in Fig.~\ref{fig:example1:state1}.

\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \begin{subcaptiongroup}
      \subcaptionlistentry{}
      \begin{minipage}{0.5\textwidth}
        \begin{tikzpicture}[outer sep=auto, level distance=4em, scale=0.50]
          \node at (-2,0) {\captiontext*{}\label{fig:example1:state1}};
          \node {$H_{y}\colon$};
          \node at (1.5,0) {2}
            child {node {$b$}
              child {node {$\{h₁\}$}}};
          \node at (5, 0) {$M_y\colon$};
          \node at (6.5, 0) {1}
            child {node {$c$}
              child {node {$\{[h₁]\}$}}};
        \end{tikzpicture}
      \end{minipage}
    \end{subcaptiongroup}
    \begin{subcaptiongroup}
      \subcaptionlistentry{}
      \begin{minipage}{0.5\textwidth}
        \begin{tikzpicture}[outer sep=auto, level distance=4em, scale=0.50]
          \node at (-2,0) {\captiontext*{}\label{fig:example1:state2}};
          \node {$H_y\colon$};
          \node at (1.5, 0) {2}
            child {node[xshift=-2] {$b$}
              child {node {$\{h₁\}$}}}
            child {node[xshift=2] {$a$}
              child {node {$\{h₂\}$}}};
          \node at (5, 0) {$M_y\colon$};
          \node at (6.5, 0) {1}
            child {node[xshift=-2] {$c$}
              child {node {$\{[h₁]\}$}}}
            child {node[xshift=2] {$b$}
              child {node {$\{[h₂]\}$}}};
        \end{tikzpicture}
      \end{minipage}
    \end{subcaptiongroup}
  \end{minipage}
  \\[1em]
  \begin{subcaptiongroup}
    \subcaptionlistentry{}
    \begin{minipage}{\textwidth}
      \begin{tikzpicture}[outer sep=auto, level distance=4em, scale=0.50]
        \node at (-2,0) {\captiontext*{}\label{fig:example2:state1}};
        \node {$H_x\colon$};
        \node at (1.5, 0) {2}
          child {node {$a$}
            child {node {$\{h_1\}$}}};
        \node at (5, 0) {$H_y\colon$};
        \node at (6.5, 0) {2}
          child {node {$a$}
            child {node {$\{h_1\}$}}};
        \node at (10, 0) {$M_x\colon$};
        \node at (11.5, 0) {1}
          child {node {$a$}
            child {node {$\{[h_2]\}$}}};
        \node at (15, 0) {$M_y\colon$};
        \node at (16.5, 0) {1}
          child {node {$b$}
            child {node {$\{[h_2]\}$}}};
      \end{tikzpicture}
    \end{minipage}
  \end{subcaptiongroup}
  \\[1em]
  \begin{subcaptiongroup}
    \subcaptionlistentry{}
    \begin{minipage}{\textwidth}
      \begin{tikzpicture}[outer sep=auto, level distance=4em, scale=0.50]
        \node at (-2,0) {\captiontext*{}\label{fig:example2:state2}};
        \node {$H_{x}\colon$};
        \node at (1.5, 0) {2}
          child {node {$a$}
            child {node {$\{h_1, h_3\}$}}};
        \node at (5, 0) {$H_y\colon$};
        \node at (6.5, 0) {2}
          child {node {$a$}
            child {node {$\{h_1\}$}}}
          child {node {$b$}
            child {node {$\{h_3\}$}}};
        \node at (10, 0) {$M_x\colon$};
        \node at (11.5, 0) {1}
          child {node {$a$}
            child {node {$\{[h_2]\}$}}};
        \node at (15, 0) {$M_y\colon$};
        \node at (16.5, 0) {1}
          child {node {$b$}
            child {node {$\{[h_2]\}$}}};
      \end{tikzpicture}
    \end{minipage}
  \end{subcaptiongroup}
  \caption{%
    Rule states encountered during Examples~1 and~2.
    The variable indices are represented as trees, with nodes in the first row denoting slots and nodes in the second row denoting variable instantiations.%
  }
\end{figure}

Now we add $h₂ : a ≤ b$, which again unifies with both slots.
For slot~1, $\sub₁(h₂)(y) = b$, so we add the match $[h₂]$ to $M_{y}(1, b)$.
We then check for hypotheses in slot~2 that also instantiate $y$ with $b$, and indeed we find $M_{y}(2, b) = \{h₁\}$.
Hence we return the complete match $[h₂, h₁]$, which induces a new hypothesis $h₃ : a ≤ c$ to be added later.

For slot~2, we add $h₂$ to $H_{y}(2, a)$.
There are no compatible matches at level~1 so we are done with $h₂$, yielding the final rule state in Fig.~\ref{fig:example1:state2}.

\paragraph{Example 2.}
We now consider a rule $r₂ : \All{x~y}{A~x~y → B~x~y → C~x~y}$ with multiple shared variables.
We add hypotheses $h₁ : B~a~a$, $h₂ : A~a~b$ and $h₃ : B~a~b$ to the rule state of $r₂$, starting with $h₁$, which is added to $H$ with slot~2.
There are no matches yet, so there is nothing to extend.
We then add $h_2 : A~a~b$, which initially generates a match $m = [h₂]$ that is added to $M$, yielding the rule state in Fig.~\ref{fig:example2:state1}.
% NOTE : uncomment to get full trees.
To find hypotheses compatible with $m$, we consider the set of variables $\vars(\lvl(m) + 1)$.
These are the variables shared between slot~$\lvl(m) + 1 = 2$ and all previous slots, i.e.\ $x$ and $y$.
Hence the compatible hypotheses are
\[
  \bigcap_{x \in \vars(\lvl(m) + 1)} H_{x}(i + 1, \sub(m)(x)) = H_{x}(2, a) \cap H_{y}(2, b) = \emptyset
\]
This means the match cannot currently be extended.
We now turn to the remaining hypothesis $h_3 : B~a~b$, adding it first to $H$.
This yields the rule state in Fig.~\ref{fig:example2:state2}.
Since $h₃$ matches slot $i = 2$, the matches that can be extended by $h₃$ are
\[
  \bigcap_{x \in \vars(i)} M_{x}(i - 1, \sub_{i}(x)) = M_{x}(1, a) \cap M_{y}(1, b) = \{[h₂]\}
\]
This produces the complete match $[h₂, h₃]$, so we can add a new hypothesis of type $C~a~b$.

\section{Implementation}%
\label{sec:implementation}

Implementing the preceding technique efficiently requires some care, particularly in the context of dependent type theory.
We briefly discuss the most prominent challenges.

\paragraph{Unification.}
In Lean's type theory, types contain programs and unification is expected to respect definitional equality, i.e.\ equality up to evaluation.
For instance, a theorem with premise $P~2$, where $P$ is some predicate, should apply to a hypothesis of type $P~((\Lam{x}{x})~2)$ since the two types are definitionally equal.
As a result, unifying premises and hypotheses is arbitrarily expensive in theory and a major cost centre in practice.
Our technique therefore seeks to minimise the number of unifications performed.

To limit the expense of unification, we perform unification at \emph{reducible} transparency.
Lean has multiple transparency levels that determine which constants are unfolded.
\enquote{Reducible} is the most restrictive of these, so we unfold few (and, in practice, only non-recursive) definitions, which greatly reduces the cost of unification.
However, this limits the power of forward reasoning since fewer terms can be unified.

We also need to account for higher-order unification problems: $Q~\mvar{f}$, where $Q$ is a predicate over functions and $\mvar{f}$ is a function variable, should unify with $Q~(\Lam{x}{x})$.
Since higher-order unification is undecidable~\cite{ThirdOrderUnificationUndecidableLucchesi,ThirdOrderUnificationUndecidableHuet,SecondOrderUnificationUndecidable}, we instead use a unification algorithm provided by Lean that is complete for first-order unification problems and heuristically solves some higher-order problems as well.
This makes our technique incomplete in the sense that some theorem instantiations will be missed, but since the same unification algorithm is used throughout Lean, users are well acquainted with this form of incompleteness.

During unification, we treat metavariables in hypotheses as constants, i.e.\ we never assign them.
This reflects the expected behaviour of forward rules: given a hypothesis $\mvar{x} < \mvar{y}$, we would not expect a forward rule to commit to the choice $\mvar{x} ≔ 1$ and $\mvar{y} ≔ 0$ just because this allows the rule to derive a contradiction.

\paragraph{Data Representation.}
Since the hypothesis index $H$ and the variable index $M$ of a rule state have the same domain, we combine them into one map.
This core data structure is then represented by nested persistent hash maps: a first hash map maps variables to a second hash map; the second hash map maps slots to a third hash map; and the third hash map maps instantiations to sets of hypotheses and sets of partial matches.

When we compare instantiations, the comparison should be up to definitional equality.
For example, when we try to apply a theorem $\All{x}{P~x → Q~x → R~x}$ where the first slot matches a hypothesis of type $P~a$ and the second slot matches a hypothesis of type $P~((\Lam{x}{x})~a)$, the application should succeed even though the instantiations of $\mvar{x}$ are different.
We therefore do not store $a$ and $(\Lam{x}{x})~a$ in the variable maps as-is.

Instead, we bring all instantiations into a normal form which we call \emph{reducible proof-irrelevant normal form} (RPINF).
This is the usual normal form of expressions with respect to Lean's notion of computation, with two modifications.
First, only \emph{reducible} defined constants are unfolded.
This reflects our choice to perform unification at reducible transparency.
Second, parts of the original expression that are proofs (in the sense that their types are members of the universe $\Prop$ of propositions) are not normalised.
This reflects the fact that Lean's type theory definitionally equates any two proofs of the same proposition, so there is no need to check whether they are equal.
When computing the RPINF, we also mark all proofs occurring in it with some special \enquote{metadata}.

With this setup, two expressions $t$ and $u$ are definitionally equal at reducible transparency if and only if their RPINFs, $t'$ and $u'$, are syntactically equal except for those portions of $t'$ and $u'$ that are proofs, which are always considered equal.
Since this is an entirely syntactic notion of equality, it can be checked quickly and we can define an effective hash function that is compatible with it.
This allows us to use a hash map for the instantiations while still comparing them up to definitional equality.

We believe that this scheme is more efficient than alternative schemes relying on repeated unification.
For example, we could use Lean's implementation of imperfect discrimination trees, which implement a map from expressions to arbitrary data where lookup respects definitional equality.
However, since these discrimination trees are imperfect, they may return spurious lookup results, so whether a lookup result is valid must be verified by unifying the query expression with the expression that was originally stored in the discrimination tree.
We believe that repeatedly performing such unifications would generally be more expensive than calculating the RPINF of each hypothesis type (which can be cached).

\paragraph{Lazy Insertion.}
Adding a hypothesis to a rule state is fairly cheap, but not free: it involves one unification for each slot that the hypothesis likely matches (according to a discrimination tree index), followed by several operations on the variable indices.
We therefore delay insertions until the index has selected at least one hypothesis for each slot.
This helps, for example, with the defining rule for involutive functions: if $f : α → α$ is involutive, then for any $a : α$ we have $f (f a) = a$.
When this rule is used as a forward rule, the premise $a$ matches any non-propositional hypothesis, but with lazy insertion we only process $a$ if there is, in fact, an involutive function is the context.

\paragraph{Match Equivalence.}
Naively, two matches $[h₁, \dots, hₙ]$ and $[h₁', \dots, hₙ']$ would be considered equal if they are structurally equal, i.e.\ $h₁ = h₁'$, \dots, $hₙ = hₙ'$.
However, consider the rule $r : \All{x~y}{P~x~y → Q~x~z → R~x}$ and the hypotheses $h₁ : P~a~b$, $h₂ : P~a~c$, $h₃ : Q~a~d$ and $h₄ : Q~a~e$.
The complete matches $m₁ ≔ [h₁, h₃]$ and $m₁' ≔ [h₁, h₄]$ are structurally different, but the difference does not matter because the applications $r~h₁~h₃$ and $r~h₁~h₄$ both have type $R~a$.
Generating both $m₁$ and $m₁'$ would therefore be redundant.
Similarly, the partial matches $m₂ ≔ [h₁]$ and $m₂' ≔ [h₂]$ are redundant because the choice between $h₁$ and $h₂$ affects neither the type of any eventual application derived from $m₂$ and $m₂'$ nor the hypotheses that can be used to complete $m₂$ and $m₂'$.
Hence we should store only one of $m₂$ and $m₂'$ in the rule state of $r$.
Based on these observations, we consider two matches $m$ and $m'$ of a rule $r$ with conclusion $A$ equal if $\lvl(m) = \lvl(m')$ and for each variable $x$ that appears either in the conclusion of $r$ or in the type of a slot $i$ of $r$ with $i > \lvl(m)$ we have $\sub(m)(x) = \sub(m')(x)$.
To ensure that we respect definitional equality, the instantiations in the substitutions are kept in RPINF.

\paragraph{Redundant Hypotheses.}
When we find a complete match, we add a new hypothesis of the corresponding type $T$ to the goal, but only if there is no other hypothesis of type $T$ already present.
To efficiently implement this check, we store in the forward match state a set containing the RPINFs of all hypotheses that have been added to (and not subsequently removed from) the state.
We modified the naive forward reasoning implementation to use the same mechanism.

\section{Evaluation}%
\label{sec:evaluation}

There are unfortunately no standard benchmarks for Aesop-like tactics, and since Aesop is not expected to solve most goals without an extensive, manually curated set of rules, benchmarks for push-button theorem provers such as TPTP~\cite{TPTP} are not very meaningful.
Additionally, forward rules are currently little used by Aesop's main client, the Mathlib mathematics library~\cite{Mathlib}.
We therefore only evaluate our technique against the previous, naive Aesop implementation on several synthetic benchmarks that demonstrate the performance characteristics of incremental forward reasoning.

For each benchmark, we set up certain forward rules and a goal with handpicked hypotheses.
We then run the \texttt{saturate} tactic, which uses either naive or incremental forward reasoning to exhaustively apply the forward rules.
Performing only forward reasoning is the best-case scenario for the incremental algorithm since previous results can be fully reused.
When forward reasoning is combined with other rules, we would expect smaller gains, depending on how much reuse the non-forward rules allow.

In what follows, instead of the natural number type of Lean's standard library, which has special support in the compiler and kernel, we use a custom type of Peano naturals.
Unifying two concrete Peano natural numbers $n$ and $m$ involves comparing $\min(n, m)$ constructors.
We can therefore use moderately large numbers to simulate the bigger unification problems that Aesop encounters, for example, in Mathlib's category theory development.

Benchmark code and data are available in our supplement~\cite{supplement}.
The results shown below were obtained on a MacBook Pro with an Apple~M2 Pro processor and \SI{32}{GB} of RAM.
The results are an average of \textcolor{red}{N} (TODO) runs per benchmark.

\paragraph{Transitivity.}
Our first benchmark is a scaled-up variant of the transitivity example from Sec.~\ref{sec:incremental}.
Given some relation $≺$ on natural numbers, we register the lemma $\All{x~y}{x ≺ y → y ≺ z → x ≺ z}$ as the sole forward rule and set up a goal with hypotheses $h₁ : a ≺ a + 1, \dots, hₙ : a + n - 1 ≺ a + n$. The test is run for $a = 0$, represented with a square \protect\showpgfsquare\, and $a = 100$, represented with a circle \protect\showpgfcircle\ .
Saturating this goal adds $n(n-1)/2$ hypotheses to the context.

\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        xlabel={Number of hypotheses},
        ylabel={Time in ms},
        xmin=1, xmax=32,
        ymin=4, ymax=2^16,
        %xtick={1,2,4,8,16},
        %ytick={2^3,2^5,2^7,2^9,2^11,2^13},
        xmode=log,
        ymode=log,
        log basis x=2,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 7.663472) (2, 9.354750) (4, 15.768194) (8, 104.921791) (16, 1882.683111) (32, 45220.919667)  
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 10.308541) (2, 10.704638) (4, 16.274305) (8, 38.800833) (16, 179.697681) (32, 1175.888750)
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 10.591138) (2, 23.881333) (4, 142.013389) (8, 1676.148375) (16, 24327.359417) (32, 392049.849708) 
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 13.974694) (2, 26.216833) (4, 68.953444) (8, 245.601028) (16, 1028.670680) (32, 4388.917500) 
          };
      \end{axis}
    \end{tikzpicture}
    \caption{Transitivity benchmark}%
    \label{fig:trans}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        xlabel={Number of Rules},
        %ylabel={Time in ms},
        xmin=1, xmax=32,
        ymin=4, ymax=2^16,
        ymode=log,xmode=log,
        log basis x=2,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]

        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 10.399958) (2, 16.426541) (4, 43.461403) (8, 205.595541) (16, 1238.090889) (32, 8927.857819)
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 11.231875) (2, 12.879500) (4, 16.851930) (8, 24.666485) (16, 42.054055) (32, 76.369778)
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 19.405222) (2, 33.487472) (4, 80.107597) (8, 284.954791) (16, 1536.089624) (32, 10721.729861)
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 25.892069) (2, 40.031930) (4, 69.446902) (8, 130.640986) (16, 260.653917) (32, 481.827902)
          };
      \end{axis}
    \end{tikzpicture}
    % graph \includegraphics[width=.4\linewidth]{image1}
    \caption{Independence benchmark}%
    \label{fig:indep}
  \end{subfigure}
  \caption{The above results depict the behavior of both algorithms in a context where a lot of work can be reused (Transitivity) and in another where no work can be reused (Independence). 
  In both benchmarks, results denoted with a square \protect\showpgfsquare\ come from a context where the argument of the predicates is small ($a = 0$), leading to easy unification problems; results denoted with a circle \protect\showpgfcircle\ come from a context where the unification problem is hard ($a = 100$).}%
  \label{fig:benchmark}
  \end{figure}

We see in Fig.~\ref{fig:trans} that the incremental algorithm has a sizable advantage over the naive algorithm.
An important detail is that the scales are logarithmic; the dotted horizontal lines represent a relative $2^3 = 8$ fold increase.
To highlight this, we consider the time taken for $2^5 = 32$ hypotheses and $a = 0$, these are the square marks on the rightmost entries: The incremental algorithm dominates the naive algorithm by a relative factor of more than 38.

These results are unsurprising:
The naive algorithm goes over the context, containing initially $n$ hypotheses, and compares all hypotheses pairwise.
This is repeated approximately for each of the $n(n-1)/2$ added hypotheses, so the overall complexity is $O(n^4)$.
The incremental algorithm considers each hypothesis only once.
The hypotheses get matched to both slots and are then compared with at most $n$ other hypotheses.
This yields an expected complexity of $O(n^2)$.

Either algorithm generates all possible transitivity chains.
If only the longest such chain, from $a$ to $a + n$, is needed, the transitivity rule can be registered as a \emph{destruct} rule, a special sort of forward rule that, whenever it is applied to certain hypotheses, removes the hypotheses from the context.
This reduces the complexity of the incremental algorithm to $O(n)$.

\paragraph{Independent Rules.}
In the last example, part of the stateful algorithm's success stems from the fact that it could reuse a lot of its previous work.
An obvious question is then how well does the stateful algorithm behave when there is les work to be reused.

Given a set of predicates $\mathcal{P} := \{P_{i,j} (x) | 1 \leq i \leq n \text{ and } 1 \leq j \leq 6\}$ and $\mathcal{Q} :=\{Q_{i}(x) | 1 \leq i \leq n \}$, in this test, we consider the $n$ following rules $r_i : \forall x, P_{i,1}(x)\to \dots P_{i,6}(x) \to Q_i(x)$.
The procedures are then run on a context containing exactly $\mathcal{P}$.
Here, as the rules are independent, the interference between them is minimal and there is less work to be reused.

This doesn't mean, however, that no work is reused: A successful rule application creates a subgoal with the newly added fact.
This means that the naive algorithm treats this subgoal as a completely new goal whereas the incremental algorithm can reuse all the work already done on the other hypotheses.
A similar analysis to the one done in the first benchmark suggests a cubic complexity in the number of rules for the naive algorithm against a linear one for the incremental algorithm.
The advantage is clearly seen in Fig.~\ref{fig:indep}.
%In this case, we sketch the behaviour of the stateless algorithm:
%It pick a first hypothesis, $h_{i,j} : P_{i,j}(a)$.
%If $j < 6$, the index returns nothing and it selects another hypothesis.
%For $h_{i,6} : P_{i,6}(a)$, the rule $r_i$ is selected, and it then needs to try and fill its remaining five slot with the context's $6n$ hypotheses which it tries all.
%This gives a complexity of about $O(n^2)$.

%For the stateful algorithm, the hypotheses are considered one by one and match directly to the rule and the slot.
%The expected complexity is then linear in the number of rules.

\paragraph{Rules of Varying Depth.}
As a last test case, we consider a scenario for which the naive algorithm should have a clear advantage over the incremental algorithm.
The main idea is the following: The naive algorithm sometimes does a better job at recognising that a rule is unprovable.
For example, when the context contains all but the last premise of a rule; the naive algorithm's index doesn't even select the rule. (See Sec. \ref{sec:naive} for the complete description of the naive algorithm.)
This leads to a much faster treatment of this rule than the incremental algorithm that will still process the hypotheses in the context that match the other premises.

Inspired by this, we find it interesting to consider a closely related test: A context with all the necessary hypotheses matching the premises however with one of them being incompatible with the others.
The reason why we introduce an incomptatible hypothesis, instead of having one absent hypothesis in the context, is because even thought the naive algorithm is sure to be fast, the incremental algorithm also handles this specific situations well.
See \textbf{Lazy Insertion} in Sec. \ref{sec:implementation}.

With all of this in mind, in Fig.~\ref{fig:erase}, the test is set up as follows: given some rule $r$ with $n$ premises $P_i(x)$, $1 \leq i \leq n$, we run the procedure in a context where $n-1$ premises are instanciated with $a$ and one of them with $a + 1$.
The numbers on the $x$ axis represent which hypothesis has the incompatible instantiation in the order of consideration by the algorithm.
In a sense, this is how \textit{deep} the naive algorithm investigate before it stops.
The position of the incompatible hypothesis has little effect on the stateful algorithm as the index selects the rule in all case.
The work the stateful algorithm has to do is thus close to constant in this scenario.
The Figure \ref{fig:erase-noprecomp} showcases that the naive algorithm has a sizable advantage in this test, especially when unification is difficult.

\paragraph{Precompilation.}

Some of the modules implemented in Lean can be precompiled.
This makes a lot of sense for our approach since it is inevitably more complex that the naive approach.
It thus benefits from getting precompiled.
This translate into even more substantial gains in the first two tests and in the third test, the incremental algorithm again takes the advantage over the naive algorithm. See \ref{fig:erase-precomp}.
%Precompilation is by default turned off for Mathlib4 as it amplifies its already long compilation time.
%Users working on projects that uses a lot of forward reasoning can however chose to pay this compilation time price once and then benefit from fast forward reasoning.

\begin{figure}
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        ylabel={Time in ms},
        xmin=1, xmax=5,
        ymin=8, ymax=2^10,
        xtick={1,2,3,4,5},
        ymode=log,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 35.466292) (2, 40.918638) (3, 46.936125) (4, 54.933124) (5, 61.269861)
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 135.967736) (2, 127.324444) (3, 136.891250) (4, 118.720444) (5, 116.293222) 
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 107.978069) (2, 115.900527) (3, 122.454958) (4, 137.732764) (5, 144.037681) 
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 619.932611) (2, 601.456805) (3, 600.926110) (4, 589.144389) (5, 595.091180) 
          };
      \end{axis}
    \end{tikzpicture}
    \caption{Depth benchmark no-precompilation}%
    \label{fig:erase-noprecomp}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \begin{tikzpicture}[scale=0.75]
      \begin{axis}[
        title={Depth Test},
        xlabel={Depth},
        %ylabel={Time in ms},
        xmin=1, xmax=5,
        ymin=8, ymax=1024,
        xtick={1,2,3,4,5},
        ymode=log,
        log basis y=2,
        legend pos=north west,
        ymajorgrids=true,
        grid style=dashed,
        legend image post style={mark=},
        ]
        \addplot[
          color=orange,
          mark=square,
          ]
          coordinates {
            (1, 13.728805) (2, 17.909194) (3, 23.324666) (4, 28.030819) (5, 33.279611) 
          };
        \addlegendentry{naive}
        \addplot[
          color=blue,
          mark=square,
          ]
          coordinates {
            (1, 13.841305) (2, 12.801791) (3, 13.358389) (4, 13.200611) (5, 13.251902)
          };
        \addlegendentry{incremental}
        \addplot[
          color=orange,
          mark=o,
          ]
          coordinates {
            (1, 81.546486) (2, 85.345444) (3, 90.532347) (4, 98.247764) (5, 105.423875) 
          };
        \addplot[
          color=blue,
          mark=o,
          ]
          coordinates {
            (1, 25.539097) (2, 24.432485) (3, 25.629666) (4, 23.935250) (5, 23.911139)
          };
      \end{axis}
    \end{tikzpicture}
    % graph \includegraphics[width=.4\linewidth]{image1}
    \caption{Depth benchmark with precompilation}%
    \label{fig:erase-precomp}
  \end{subfigure}
  \caption{The above results depict the behavior of both algorithms in depending on the \text{depth} of potential partial matches. The lines with square entry points represent problems where unification is easy. The lines with circle entry points are the same tests with hard unification problems.}%
  \label{fig:erase}
  \end{figure}

\section{Related Work}

\paragraph{ACL2 and PVS.}
Among white-box proof automation tactics, to our knowledge only ACL2's \enquote{waterfall}~\cite{ACL2} and PVS's various proof search strategies~\cite{ACL2,PVS} have special support for forward reasoning.
Both use essentially Aesop's naive algorithm, i.e.\ the context of each goal is saturated independently without reusing partial results from other goals.
ACL2 additionally supports \emph{trigger terms}, which allow forward rules to be selected based on the presence of a particular term anywhere in the goal, rather than based on particular hypotheses.
Aesop implements a similar mechanism, \emph{rule patterns}, not described here due to space constraints.
Some other white-box tactics, e.g.\ Rocq's auto, can execute arbitrary tactics including forward rules, but make no special provisions for them.

\paragraph{SATCHMO.}
The SATCHMO first-order theorem prover~\cite{SATCHMO} uses a search procedure that treats a given set of implications as forward rules and uses a single non-forward rule to perform case analysis on disjunctive hypotheses.
SATCHMO-style provers may benefit from our incremental approach if the set of implications is large.
However, since unification is comparatively cheap in first-order logic, it is not clear whether avoiding repeated unifications outweighs the cost of maintaining the forward state.
SATCHMO also employs an optimisation that \enquote*{focuses} forward reasoning on the hypothesis established by the last case split, providing a different form of incrementality.

\paragraph{Knowledge-Based Systems.}
Some knowledge-based reasoning systems, e.g.\ Algernon~\cite{Algernon} and EYE~\cite{EYE}, support combined forward and backward reasoning.
However, to our knowledge, no such system allows non-forward rules to change the assumptions available to forward rules, so they do not require incrementality.
Nevertheless, our technique bears some resemblance to the RETE algorithm~\cite{RETE} for forward reasoning in large knowledge bases.

\jcom{auto2?}

\section{Conclusion}

We have presented a technique that integrates forward reasoning into a tree-based proof search procedure supporting arbitrary proof rules.
We avoid redundant reasoning steps by caching partial forward rule applications in a data structure, the \emph{forward match state}, that enables efficient updates.
The technique is independent of the search strategy and logic used.
Our evaluation shows substantial speedups on some synthetic benchmarks, relative to Aesop's previous implementation.

\bibliography{lit}

\end{document}